```{r}
library(knitr)
```
---
Title: "PA1_template"
author: "Anusha Verankki"
date: "July 2, 2017"
output: html_document
==================================================================================
Loading the data into a dataframe
```{r read_data, echo=TRUE}
act.dat<- read.csv("activity.csv")
```

Tidying the data set by omitting all the NA values and storing in new dataframe
```{r tidy_data, echo = TRUE}
actdat.tidy<- na.omit(act.dat)
#Part 1: Finding the total number of steps taken in a day
q1.sum<- aggregate(steps~date,actdat.tidy,sum)

```

Creating a histogram of the sum of each steps taken in a day, per interval
```{r Plot1, echo = TRUE}
#Plotting the histogram of total number of steps taken in a day 
hist(q1.sum$steps, xlab = "Frequency", ylab = "Number of steps in a day", main = "Plot 1: Total steps taken per day")
```

finding the mean and median values in the tidy data set
```{r mean&median, echo=TRUE}
q1.mean<- mean(q1.sum$steps)
q1.median<- median(q1.sum$steps)
print(paste0("Mean is ",round(q1.mean,2)))
print(paste0("Median is ",round(q1.median,2)))
  
#Finding the average number of steps taken in a day during an interval
q2avg<- with(actdat.tidy,tapply(steps,interval,mean))
intervals<- actdat.tidy$interval
q2mean<-as.data.frame(cbind(intervals,q2avg))

```
Plotting the average number of steps taken in a day per interval
```{r Plot2, echo=TRUE}
plot(q2mean$intervals, q2mean$q2avg, type = "l", xlab = "Intervals", ylab = "average steps", main = "Plot 2: Avg number of steps per interval")
```

Finding the interval with the maximum steps
```{r max_interval, echo=TRUE}
#Finding the Max steps
maxsteps<- max(q2avg)
#Finding the interval with the max steps
max.int<- q2mean[(q2mean$q2avg==maxsteps),]$intervals
```

Filling in the NA values in the data frame
```{r filling_NA, echo=TRUE}

# finding the total number of NAs 
sum.na<- sum(is.na(act.dat))
print(paste0("Total number of NAs is ", sum.na))

#Top 5 rows of dataset before filling in NA values
head(act.dat)

#filling in missing values with average steps of that day
k1<- dim(act.dat)
q1.means<- aggregate(steps~date,act.dat, mean, na.action = NULL)
avg.steps<- mean(act.dat$steps, na.rm = TRUE)

#if the average steps for the day does not exist, insert the average steps for the dataset

q1.means[is.na(q1.means)]<- avg.steps
k2<- dim(q1.means)

for (x in 1:k1[1]){
    temp1<- act.dat[x,1]
    
    for (y in 1:k2[1]){
      if ((is.na(temp1))&&(act.dat[x,2]==q1.means[y,1])) act.dat[x,1]<- q1.means[y,2]
    }
}

#Top 5 rows of the dataset after filling in NA values
head(act.dat)
```

Creating a histogram of the average steps per interval
```{r Plot3, echo=TRUE}
# histogram of total number of steps taken per day
sum_updated<- aggregate(steps~date,act.dat,sum)
hist(sum_updated$steps, xlab = "Frequency", ylab = "Number of steps in a day", main = "Plot 3: Total steps taken")
```
Recalculating the mean and median
```{r mean&median2, echo=TRUE}
q1.mean.updated<- mean(sum_updated$steps)
q1.median.updated <- median(sum_updated$steps)
print(paste0("Mean is ",round(q1.mean.updated,2)))
print(paste0("Median is ",round(q1.median.updated,2)))

```

Examining the data on weekdays vs weekends 

```{r weekday_data, echo=TRUE}
daycol<- vector()
wd<- weekdays(as.Date(act.dat$date))

# adding a new column named daycol with 2 levels: weekday/weekend
for (i in 1:k1[1]){
  daycol[i]<- if (wd[i]=="Sunday") "weekend" else if (wd[i]=="Saturday") "weekend" else "weekday"
}
act.dat<- cbind(act.dat,daycol)
```
Displaying results on weekends vs weekdays 
```{r Plot4, echo = TRUE}
library(ggplot2)
activity_day<- aggregate(steps~interval + daycol, act.dat, mean)
dayplot<- ggplot(activity_day, aes(x = interval, y = steps, color = daycol))+geom_line()+labs (x = "Interval", y = "avg steps", title = "Plot 4: Avg steps per interval(weekdays vs weekdays)")
print(dayplot)
```
